---
title: "Mastering Shiny: Shiny modules (apps)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mastering Shiny: Shiny modules (apps)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
#| eval: true
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 3.5,
  dpi = 140,
  fig.retina = TRUE
)
```

```{r}
#| label: pkgs
#| message: false
#| warning: false
library(msst2ap)
library(lobstr)
library(fs)
```

## Mastering Shiny Apps


This vignette walks through the structure of each application using the abstract syntax tree (`ast()`) function from the [`lobstr` package](https://lobstr.r-lib.org/). Three of the applications in the `msst2ap` package come from the [Modules section of Mastering Shiny](https://mastering-shiny.org/scaling-modules.html), with a few adjustments: 

-   Each standalone app function has a `verbatimTextOutput()` in the `ui`: 

    ```{r}
    #| eval: false
    #| echo: true
    shiny::verbatimTextOutput("vals")
    ```

-   And an accompanying `renderPrint()` in the `server`:

    ```{r}
    #| eval: false
    #| echo: true
    output$vals <- shiny::renderPrint({
      x <- shiny::reactiveValuesToList(input,
                              all.names = TRUE)
      print(x)
    })
    ```


Printing the output from `shiny::reactiveValuesToList()` in the top-level of the standalone function shows the namespaces and modules throughout the application. 


### `datasetApp()`
    
```{verbatim}
#| eval: false 
#| echo: true
█─datasetApp 
├─filter = NULL 
└─█─shinyApp 
  ├─ui = █─fluidPage 
  │      ├─█─datasetInput 
  │      │ ├─id = "dataset" 
  │      │ ├─filter = is.data.frame 
  │      │ └─█─selectInput 
  │      │   └─inputId = "dataset" 
  │      └─█─tableOutput 
  │        └─"data"
  └─server = █─function(input, output, session)
             └─█─data <- datasetServer(id = "dataset") 
               └─output$data <- renderTable(head(data()))
```

**`datasetApp()` contains the `datasetInput()` and `datasetServer()` module functions**

-   `datasetInput()` displays the objects from the `datasets` package (`"dataset"`) matching the `filter` argument (i.e., `is.data.frame`, `is.matrix`) and the table output (`"data"`) in the UI
   
-   The `inputId` in the UI passes a `datasets` object (as `input$dataset`, a character string) to `datasetServer()`, which fetches and returns the object as a reactive via `get()` 

#### `datasetApp()` Reactive values 

The output from `reactiveValuesToList()` are displayed in the UI: 

```
$`dataset-dataset`
[1] "ability.cov"
```

The first `dataset` is from the module `id`, and the second is from the `inputId` (in the `selectInput()`).

<!--
    
```{r}
#| echo: false 
#| eval: false
#| comment: ""
#| include: true
lobstr::ast(
  datasetApp(filter = NULL,
    shinyApp(
      ui = fluidPage(
            datasetInput(id = "dataset", filter = filter,
              selectInput(inputId = "dataset")
              ),
            tableOutput("data")),
  server = function(`input, output, session`) {
    `data <- datasetServer(id = "dataset")`
    `output$data <- renderTable(head(data()))`
    }
   )
  )
)
```

-->
    
### `selectVarApp()`

**`selectVarApp()` contains the `datasetInput()`/`Server()` and `selectVarInput()`/`Server()` modules**


```{verbatim}
#| eval: false 
#| echo: true
█─selectVarApp 
└─█─shinyApp 
  ├─ui = █─fluidPage 
  │      ├─█─datasetInput 
  │      │ ├─id = "data" 
  │      │ ├─filter = is.data.frame
  │      │ └─█─selectInput 
  │      │   └─inputId = "dataset" 
  │      ├─█─selectVarInput 
  │      │ ├─id = "var" 
  │      │ └─█─selectInput 
  │      │   └─inputId = "var" 
  │      └─█─tableOutput 
  │        └─"out" 
  └─server = █─function(input, output, session)
             └─█─data <- datasetServer("data")
               █─var <- selectVarServer("var", data, filter = is.numeric)
               └─output$out <- renderTable(head(var()))
             
```

- **`dataset` module:**
    -   [x] `datasetInput("data")` 
  
    -   [x] `datasetServer("data")`
    
- **`selectVar` module:**
      
  -   `selectVarInput()` displays the variables (columns) in the returned `datasets` object from `datasetServer()` (as `"var"`) and the rendered output (as `"out"`)    
  -   In `selectVarServer()`, the columns in the `data` returned from `datasetServer()` is filtered to those columns matching the `filter` argument (i.e., `is.numeric`), and the selected `"var"` is displayed in `"out"`

### `selectVarApp()` Adjustments 

The variable value from `selectVarServer()` was originally returned with a simple call to `reactive()`

```{r}
#| eval: false 
#| echo: true
selectVarServer <- function(id, data, filter = is.numeric) {
  stopifnot(is.reactive(data))
  stopifnot(!is.reactive(filter))
  
  moduleServer(id, function(input, output, session) {
    observeEvent(data(), {
      updateSelectInput(
        session = session,
        inputId = "var", 
        choices = find_vars(data = data(), filter = filter))
    })
    
    reactive(data()[[input$var]]) # <- simple reactive()
    
  })
}
```

I've updated this with some control flow statements to ensure the variable is contained the data object. I've also changed the returned object to be a single column `data.frame` (as opposed to a vector):

```{r}
#| eval: false 
#| echo: true
selectVarServer <- function(id, data, filter = is.numeric) {
  stopifnot(shiny::is.reactive(data))
  stopifnot(!shiny::is.reactive(filter))

  shiny::moduleServer(id, function(input, output, session) {

    shiny::observe({
      shiny::updateSelectInput(
        session, "var",
        choices = find_vars(data(), filter))
    }) |>
      shiny::bindEvent(data())

    return(
      shiny::reactive({
        if (input$var %in% names(data())) { # <- updated with control flow
          data()[input$var] # returned as a single column
        } else {
          NULL
        }
      }) |>
      shiny::bindEvent(input$var)
    )

  })
}
```


#### `selectVarApp()` Reactive values 

The `reactiveValuesToList()` output for both modules are below:

```
$`var-var`
[1] "Ozone"

$`data-dataset`
[1] "airquality"
```

- The first `"var"` is from the shared `id` between `selectVarInput()`/`selectVarServer()`, the second is from the `inputId` in the `selectInput()`.

- `"data"` is from the shared `id` between `datasetInput()`/`datasetServer()`, and the second is from the `selectInput()`.

<!--

```{r}
#| echo: false 
#| eval: false
#| comment: ""
#| include: true
lobstr::ast(
  selectVarApp(
    shinyApp(
      ui = fluidPage(

      datasetInput(id = "data", filter = is.data.frame),

      selectVarInput(id = "var"),

      verbatimTextOutput("out")
        
  ), 
      server = function(`input, output, session`) {
        
        `data <- datasetServer("data")`
        
        `var <- selectVarServer("var", data, filter = filter)`

        `output$out <- renderPrint(head(var()))`
        
        }
      )
    )
  )
```

-->

### `selectDataVarApp()`
      

```{verbatim}
#| eval: false 
#| echo: true
█─selectDataVarApp 
└─█─shinyApp 
  ├─ui = █─fluidPage 
  │      └─█─sidebarLayout 
  │        ├─█─sidebarPanel 
  │        │ └─█─selectDataVarUI 
  │        │   ├─"var" 
  │        │   ├─█─datasetInput 
  │        │   │ ├─█─NS 
  │        │   │ │ ├─id 
  │        │   │ │ └─"data" 
  │        │   │ └─filter = is.data.frame 
  │        │   └─█─selectVarInput 
  │        │     └─█─NS 
  │        │       ├─id 
  │        │       └─"var" 
  │        └─█─mainPanel 
  │          └─█─tableOutput 
  │            └─"out" 
  └─server = █─function(input, output, session)
             ├─█─var <- selectDataVarServer(id = "var", filter = is.numeric)
             │ ├─data <- datasetServer(id = "data")
             │ ├─var <- selectVarServer(id = "var", data, filter = filter) 
             │ └─var 
             └─output$out <- renderTable(head(var()))
```

**`selectDataVarApp()` nests the `dataset` and `selectVar` modules *inside* the `selectDataVarUI()` and `selectDataVarServer()` functions:**

Inside `selectDataVarUI()`:

  - **`datasetInput("data")`:**
      -   [x] `datasetInput()` is moved into the `sidebarPanel()` to display the objects from the `datasets` package (`"dataset"`) matching the `filter` argument (i.e., `is.data.frame`, `is.matrix`) 
  - **`selectVarInput("var")`:**
      -   [x] `selectVarInput("var")` also displays the variables in the `sidebarPanel()` from the returned `data` object in `datasetServer()`
  
Inside `selectDataVarServer()`:

  - **`datasetServer()`:**
      -   [x] The `inputId` in the UI passes a `datasets` object (as `input$dataset`, a character string) to `datasetServer()`, which fetches and returns the object as a reactive with `get()`
  - **`selectVarServer()`:**
      -   [x] In `selectVarServer()`, the columns in the `data` returned from `datasetServer()` are filtered to only those columns matching the `filter` argument (i.e., `is.numeric`), and the selected `"var"` is returned as `var`
      -   [x] The `var()` reactive output is rendered in the UI

#### `selectDataVarApp()` Reactive values

By placing the `datasetInput()` and `selectVarInput()` functions *inside* `selectDataVar()` with another call to `NS()`, these functions end up with **two** appended namespaces: 

```
$`var-var-var`
[1] "Ozone"

$`var-data-dataset`
[1] "airquality"
```

<!--

```{r}
#| echo: false 
#| eval: false
#| comment: ""
#| include: true
lobstr::ast(
selectDataVarApp(
  shinyApp(
    ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        selectDataVarUI("var",
          datasetInput(NS(id, "data"), filter = is.data.frame),
          selectVarInput(NS(id, "var"))
          )
        ),
      mainPanel(
        verbatimTextOutput("out")
        )
    )
  ),
    
  server = function(`input, output, session`) {

    `var <- selectDataVarServer("var", filter)`(`data <- datasetServer("data")`,
      `var <- selectVarServer("var", data, filter = filter)`, `var`)
    `output$out <- renderPrint(head(var()))`

  }
  )
  ))
```
-->

### `histogramApp()`


```{verbatim}
#| eval: false 
#| echo: true
█─histogramApp 
└─█─shinyApp 
  ├─ui = █─fluidPage 
  │      ├─█─sidebarLayout 
  │      │ └─█─sidebarPanel 
  │      │   ├─█─datasetInput 
  │      │   │ ├─"data" 
  │      │   │ └─filter = is.data.frame 
  │      │   └─█─selectVarInput 
  │      │     └─"var" 
  │      └─█─mainPanel 
  │        └─█─histogramOutput 
  │          └─"hist" 
  └─server = █─function(input, output, session)
             ├─data <- datasetServer('data')
             ├─x <- selectVarServer('var', data)
             ├─histogramServer('hist', x) 
             └─output$out <- renderPrint(head(var()))
```

This [`histogramApp()`](https://mastering-shiny.org/scaling-modules.html#case-study-histogram) function has been slightly altered to accommodate converting the graph output to use `ggplot2`.


### Adjustments 

The original `histogramServer()` function uses `shiny::req()` to ensure the `x()` reactive is numeric (`is.numeric()`). It then passes this numeric vector directly to `hist()` (along with the `breaks` and title (in `main`)).

```{r}
#| eval: false 
#| echo: true
histogramServer <- function(id, x, title = reactive("Histogram")) {
  stopifnot(is.reactive(x))
  stopifnot(is.reactive(title))
  
  moduleServer(id, function(input, output, session) {
    output$hist <- renderPlot({
      req(is.numeric(x()))
      main <- paste0(title(), " [", input$bins, "]")
      hist(x(), breaks = input$bins, main = main)
    }, res = 96)
  })
}
```

I've changed `histogramServer()` slightly: I removed the `is.numeric()` function from `shiny::req()` and use `purrr::as_vector()` to convert the single column `data.frame` to a vector:  

```{r}
#| eval: false 
#| echo: true
histogramServer <- function(id, x, title = reactive("Histogram")) {
  stopifnot(shiny::is.reactive(x))
  stopifnot(shiny::is.reactive(title))

  shiny::moduleServer(id, function(input, output, session) {

    output$hist <- shiny::renderPlot({
      shiny::req(x()) # require x() (but not numeric) 
      main <- paste0(title(), " [bins =", input$bins, "]")
      hist(
        purrr::as_vector(x()), # convert the single column data.frame to a vector
        breaks = input$bins,
        main = main)
    }, res = 96)

  })
}
```

#### Reactive values

The output from `reactiveValuesToList()` is below:

```
$`hist-bins`
[1] 10

$`var-var`
[1] "Ozone"

$`data-dataset`
[1] "airquality"
```

The reason all of these modules are interchangeable is that none of them deal with rendering outputs--all of the `*Output()`/`render()*` functions are kept in the standalone application functions. 


<!--

```{r}
#| echo: false 
#| eval: false
#| comment: ""
#| include: true
lobstr::ast(
histogramApp(
  shinyApp(
    ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        datasetInput("data", filter = is.data.frame),
          selectVarInput("var")
          )
        ),
      mainPanel(
        histogramOutput("hist")
        )
    ),
  server = `function(input, output, session)`
    (
      `data <- datasetServer('data')`, 
      `x <- selectVarServer('var', data)`,
      `histogramServer('hist', x)`,
      `output$out <- renderPrint(head(var()))`
      )
  )
  )
  )
```

-->
