---
title: "shinytest2 setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{shinytest2 setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: packages 
#| warning: false 
#| message: false 
library(msst2ap)
library(shinytest2)
```

This vignette walks through setting up `shinytest2` for a shiny app-package.


## Set up

[`shinytest2`](https://rstudio.github.io/shinytest2/index.html) requires a few steps to get up and running, so I've included my start up steps below (in hopes that someone might find them useful).

## Chromium

First make sure you have [Chromium headless browser](https://www.chromium.org/chromium-projects/) installed. Chromium is the browser used to test and debug shiny apps with `shinytest2`.

If you're using macOS, you can install Chromium using [homebrew](https://formulae.brew.sh/cask/chromium): 

```{bash}
#| eval: false
#| code-fold: false 
brew install --cask chromium
```

```{bash}
#| eval: false
==> Downloading https://download-chromium.appspot.com/dl/Mac?type=snapshots
==> Downloading from https://commondatastorage.googleapis.com/chromium-browser-s
######################################################################### 100.0%
Warning: No checksum defined for cask 'chromium', skipping verification.
==> Installing Cask chromium
==> Moving App 'Chromium.app' to '/Applications/Chromium.app'
==> Linking Binary 'chromium.wrapper.sh' to '/usr/local/bin/chromium'
🍺  chromium was successfully installed!
```


## `chromote`

The [`chromote` package](https://rstudio.github.io/chromote/index.html) allows R to open Chromium. I had to make sure R knew where to find the Chromium application ([this section helps](https://rstudio.github.io/chromote/#specifying-which-browser-to-use)) by passing the path to the app to `Sys.setenv()` (*you can locate the path to Chromium using by holding <kbd>command (⌘)</kbd> and clicking on the application icon*)

```{r}
#| eval: false
remotes::install_github("rstudio/chromote", force = TRUE, quiet = TRUE)
library(chromote)
```

Find your `Chromium` app: 

```{r}
#| eval: true
chromote::find_chrome()
```

This didn't work for me (kept locating Chrome.app), so I had to find it with the Terminal:

```{r}
Sys.setenv(CHROMOTE_CHROME = "/Applications/Chromium.app/Contents/MacOS")
```

I also verified the Chromium app opened with `ChromoteSession$new()`

```{r}
#| eval: false
b <- ChromoteSession$new()
b$view()
```



