---
title: "testServer tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testServer tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
#| eval: true
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: pkgs
#| message: false
#| warning: false
library(msst2ap)
library(testthat)
```

The `testServer()` tests have been removed from `msst2ap` to focus on the `shinytest2` tests, but I've stored each each of them in this vignette for completeness. 

```{r}
#| label: test_cmt 
#| include: false
test_cmt <- function(test, msg) {
  cat("\n\t", glue::glue("  {test}: {msg}"), "\n")
}

```

## datasetServer() 

Verify `NULL` 

```{r}
shiny::testServer(datasetServer, expr = {

  testthat::expect_equal(input$dataset, NULL)
  test_cmt("datasetServer", "dataset$input is NULL")
})
```

Verify input 

```{r}
shiny::testServer(datasetServer, expr = {
  session$setInputs(dataset = "faithful")
  testthat::expect_equal(
    object = input$dataset,
    expected = "faithful")
  test_cmt("datasetServer", "dataset$input")
})
```

Verify `class()` of `returned()`

```{r}
class(datasets::airquality)
```


```{r}
shiny::testServer(datasetServer, expr = {
  session$setInputs(dataset = "airquality")
  testthat::expect_equal(
    object = class(session$returned()),
    expected = "data.frame")
  test_cmt("datasetServer", "class(session$returned())")
})
```

Verify `class()` of `returned()`


```{r}
is.matrix(datasets::WorldPhones)
```

```{r}
shiny::testServer(datasetServer, expr = {
  session$setInputs(dataset = "WorldPhones")
  testthat::expect_true(
    object = is.matrix(session$returned()))
  test_cmt("datasetServer", "is.matrix(session$returned())")
})
```

Verify `typeof()` of `returned()`

```{r}
typeof(datasets::mtcars)
```


```{r}
shiny::testServer(datasetServer, expr = {
  session$setInputs(dataset = "mtcars")
  expect_equal(
    object = typeof(session$returned()),
    expected = typeof(datasets::mtcars))
  test_cmt("datasetServer", "typeof(session$returned())")
})
```

## datasetApp()

Check `dataset` input

```{r}
ds_app <- datasetApp()
shiny::testServer(app = ds_app, expr = {
  session$setInputs(`dataset-dataset` = "chickwts")
  testthat::expect_equal(
    object = input$`dataset-dataset`,
    expected = "chickwts")
  test_cmt("datasetApp", "input$`dataset-dataset`")
})
```

Check `class` of reactive `data()`

```{r}
ds_app <- datasetApp()
shiny::testServer(app = ds_app, expr = {
  session$setInputs(`dataset-dataset` = "chickwts")
  testthat::expect_true(
    object = is.data.frame(data()))
  test_cmt("datasetApp", "is.data.frame(data())")
})
```

```{r}
ds_app <- datasetApp()
shiny::testServer(app = ds_app, expr = {
  session$setInputs(`dataset-dataset` = "chickwts")
  testthat::expect_equal(
    object = names(data()),
    expected = names(datasets::chickwts))
  test_cmt("datasetApp", "names(data())")
})
```


```{r}
ds_app <- datasetApp()
shiny::testServer(app = ds_app, expr = {
  session$setInputs(`dataset-dataset` = "chickwts")
  testthat::expect_equal(
    object = class(output$data),
    expected = "character")
  test_cmt("datasetApp", "class(output$data)")

})
```

## selectVarServer() 

The `selectVarServer()` function has additional arguments, so the `args` list is used (with `reactive()`):

```{r}
shiny::testServer(selectVarServer,

  args = list(data = reactive(datasets::mtcars),
              filter = is.numeric), expr = {

  testthat::expect_true(
    object = is.reactive(data))
  test_cmt("selectVarServer", "is.reactive(data())")
})
```

Test passing different arguments to `find_vars()`

```{r}
shiny::testServer(selectVarServer,
  args = list(data = reactive(datasets::chickwts),
              filter = is.numeric), expr = {
  testthat::expect_equal(
    object = find_vars(data(), is.factor),
    expected = "feed")
  test_cmt("selectVarServer", "find_vars()")
})
```

Test passing `filter` and setting `input$var`:

```{r}
shiny::testServer(selectVarServer,
  args = list(data = reactive(datasets::chickwts),
              filter = is.numeric), expr = {

  session$setInputs(var = "weight")
  testthat::expect_equal(object = input$var,
      expected = "weight")
  test_cmt("selectVarServer", "input$var")
})
```

Test passing `filter`, setting `input$var`, and `session$returned()`:

```{r}
shiny::testServer(selectVarServer,
  args = list(data = reactive(datasets::chickwts),
              filter = is.numeric), expr = {
  session$setInputs(var = "feed")
  testthat::expect_equal(object = session$returned(),
      expected = datasets::chickwts["feed"])
  test_cmt("selectVarServer", "session$returned()")

})
```


## selectVarApp()

Test `is.reactive(var)`

```{r}
sv_app <- selectVarApp()
shiny::testServer(app = sv_app, expr = {
  session$setInputs(`var-var` = "Ozone",
                    `data-dataset` = "airquality")
  
  testthat::expect_true(object = is.reactive(var))
  test_cmt("selectVarApp", "is.reactive(var)")
})
```

Test the inputs with module namespaces 

```{r}
sv_app <- selectVarApp()
shiny::testServer(app = sv_app, expr = {
  session$setInputs(`var-var` = "Ozone",
                    `data-dataset` = "airquality")

  testthat::expect_equal(
    object = input$`var-var`,
    expected = "Ozone")
  test_cmt("selectVarApp", "input$`var-var`")

})
```

Test `data()` is reactive

```{r}
sv_app <- selectVarApp()
shiny::testServer(app = sv_app, expr = {
  session$setInputs(`var-var` = "Ozone",
                    `data-dataset` = "airquality")

  testthat::expect_true(object = is.reactive(data))
  test_cmt("selectVarApp", "is.reactive(data)")

})
```

Test `class()` of `data()`

```{r}
sv_app <- selectVarApp()
shiny::testServer(app = sv_app, expr = {
  session$setInputs(`var-var` = "Ozone",
                    `data-dataset` = "airquality")
  
  testthat::expect_true(
    object = is.data.frame(data()))
  test_cmt("selectVarApp", "is.data.frame(data())")

})
```

Test specific variables from `data()`

```{r}
sv_app <- selectVarApp()
shiny::testServer(app = sv_app, expr = {
  session$setInputs(`var-var` = "Ozone",
                    `data-dataset` = "airquality")

  testthat::expect_equal(
    object = data()[[input$`var-var`]],
    expected = airquality[["Ozone"]])
  test_cmt("selectVarApp", "data()[[input$`var-var`]]")

})
```



## selectDataVarServer()

Test `is.reactive()` on `data`

```{r}
shiny::testServer(selectDataVarServer, 
  args = list(filter = is.numeric), expr = {
    
  testthat::expect_true(
    object = is.reactive(data))
  test_cmt("selectDataVarServer", "is.reactive(data)")

})
```

Test `is.reactive()` on `var`

```{r}
shiny::testServer(selectDataVarServer, 
  args = list(filter = is.numeric), expr = {

  testthat::expect_true(
    object = is.reactive(var))
  test_cmt("selectDataVarServer", "is.reactive(var)")

})
```


## selectDataVarApp()

Test inputs for `var-var-var`

```{r}
dv_app <- selectDataVarApp()
shiny::testServer(app = dv_app, expr = {
  session$setInputs(`var-var-var` = "Ozone",
                    `var-data-dataset` = "airquality")

  testthat::expect_equal(
    object = input$`var-var-var`,
    expected = "Ozone")
  test_cmt("selectDataVarApp", "input$`var-var-var`")

})
```

Test inputs for `var-data-dataset`

```{r}
dv_app <- selectDataVarApp()
shiny::testServer(app = dv_app, expr = {
  session$setInputs(`var-var-var` = "Ozone",
                    `var-data-dataset` = "airquality")

  testthat::expect_equal(
    object = input$`var-data-dataset`,
    expected = "airquality")
  test_cmt("selectDataVarApp", "input$`var-data-dataset`")

})
```

Test `is.reactive()` for `var`

```{r}
dv_app <- selectDataVarApp()
shiny::testServer(app = dv_app, expr = {
  session$setInputs(`var-var-var` = "Ozone",
                    `var-data-dataset` = "airquality")

  testthat::expect_true(object = is.reactive(var))
  test_cmt("selectDataVarApp", "is.reactive(var)")

})
```
