---
title: "Unit tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unit tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
#| eval: true
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 3.5,
  dpi = 140,
  fig.retina = TRUE
)
```

```{r}
#| label: pkgs
#| message: false
#| warning: false
library(msst2ap)
library(testthat)

```

## Unit tests

The unit tests have been removed from `msst2ap` to focus on the `shinytest2` tests, but I've stored each each of them are stored the tests for each R file in this vignette for completeness. 

```{r}
#| label: test_cmt 
#| include: false
test_cmt <- function(test, msg) {
  cat("\n\t", glue::glue("  {test}: {msg}"), "\n")
}

```

## find_vars()

Test function (`is.double`)

```{r}
testthat::test_that("find_vars", {
    testthat::expect_equal(
    object = find_vars(datasets::airquality,
                            filter = is.double),
    expected = names(airquality)[vapply(airquality, is.double, logical(1))])
    test_cmt("find_vars", "is.double()")
})
```

Test function (`is.character`)

```{r}
testthat::test_that("find_vars", {
  testthat::expect_equal(
    object = find_vars(datasets::mtcars,
                            filter = is.character),
    expected = datasets::mtcars |>
      dplyr::select(where(is.character)) |>
      names())
  test_cmt("find_vars", "is.character()")
})
```

Test error: `state.abb` is a vector--pass `is.character` function to `filter`

```{r}
#| error: true
find_vars(data = datasets::state.abb, is.character)
```


```{r}
testthat::test_that("find_vars", {
  testthat::expect_error(object =
      find_vars(datasets::state.abb, filter = is.character))
  test_cmt("find_vars", "error: data")
})
```

Test error: pass `is.miracle` function to `filter`

```{r}
#| error: true
find_vars(data = datasets::mtcars, is.miracle)
```


```{r}
testthat::test_that("find_vars", {
  testthat::expect_error(object =
      find_vars(datasets::mtcars, filter = is.miracle))
  test_cmt("find_vars", "error: filter")

})
```

